---
title: "FGCZ Two-Group Analysis for: `r grp2$projectName` \n Statistics for a Quantitative Protein Matrix"
author: "Functional Genomics Center Zurich"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
pdf_document:
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

if(!exists("progress")){
  progress <- function(howmuch, detail){
  }
}
```

# Input Matrix

Experiment is called: `r grp2$experimentName`

## Numbers

The protein matrix is filtered like this:

- Minimum number of peptides / protein: `r grp2$nrPeptides`
- Maximum of missing values per protein : `r grp2$maxNA`

The total number of proteins in this experiment is: `r nrow(grp2$proteinIntensity)`


# The Groups


```{r}

library(knitr)

tab <- data.frame(table(grp2$getAnnotation()$Condition))
colnames(tab) <- c("Condition","# samples")
knitr::kable(tab)

```

Here the files in each group:

```{r }
tab<-grp2$getAnnotation()[,c("Condition","Raw.file")]
rownames(tab) <- NULL
knitr::kable(tab)

```

\pagebreak

# Proteins Used for Quantitation


```{r histmissing }
missing <- grp2$getNrNAs()
int <- apply(grp2$proteinIntensity,1,sum, na.rm=TRUE)
grp2$proteinIntensity <- grp2$proteinIntensity[order(missing, -int,decreasing = T),]

```


The input matrix has the following structure.


```{r overview, fig.with=9,fig.height=6,dpi=300,fig.cap="Heatmap for quantifyable proteins sorted by missigness and intensity (log2). Gray - missing values"}
library(quantable)

imageWithLabels(t(log2(grp2$proteinIntensity)), col.labels = NULL, col=quantable::getBlueScale(21))

```

```{r, fig.cap="# of missing values per protein"}
plot(table(missing))

progress(0.1, "Summary")

```

\pagebreak


# Distribution of intensities

Shown in Figure \@ref(distributionRaw) are the un-normalized values while in Figure  \@ref(normalized)
are the z transformed values (subtracted median and divided by variance).

```{r distributionRaw, fig.width=7, fig.height=7, fig.cap="Density plot for quantifyable proteins (log2 transformed)"}
library(ggplot2)
library(reshape2)
longm <- melt(log2(grp2$proteinIntensity))

p <- qplot( variable , value , data=longm , geom="violin" , xlab="" , ylab="log2(I)")
p + stat_summary(fun.y=median,geom='point') +theme(axis.text.x = element_text(angle = 90, hjust = 1))

```



```{r normalized,  fig.width=7,fig.height=7,dpi=300, fig.cap="Density plot for normalized values (log2)"}

longm <- melt(grp2$getNormalized()$data)
p <- qplot( variable , value , data=longm , geom="violin" , xlab="" , ylab="log2(I)")
p + stat_summary(fun.y=median,geom='point') + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

The scaling factors are visualized in Figure \@ref(scaling). It shows with what factor the individually samples
are normalized. Large differences in the scaling factors are critical.


```{r  scaling,dpi=300, fig.height=6, fig.cap="Applied scaling factors for normalization (calculated using median normalization)."}
par(mfrow=c(2,1))
par(mar=c(3,10,3,3))

barplot(grp2$getNormalized()$medians,horiz=T,las=2, main="median correction", cex.names = 0.6)
barplot(grp2$getNormalized()$mads,horiz=T, las=2 , main="standard deviation correction", cex.names = 0.6)

progress(0.2, "Normalization")

```


\pagebreak

# Coefficients of variation

```{r CViolin, fig.cap="distribution of protein CV within condition and overall"}
cond1 <- grp2$getConditionData(grp2$conditions[1])
cond2 <- grp2$getConditionData(grp2$conditions[2])



cond1 <- quantable::CV(cond1)
cond2 <- quantable::CV(cond2)
all <- quantable::CV(grp2$proteinIntensity)
CVs <- rbind(data.frame(condition=grp2$conditions[1], cv=cond1),
             data.frame(condition=grp2$conditions[2],cv=cond2 ),
             data.frame(condition="all", cv=all))

p <- qplot( condition , cv , data=CVs , geom="violin" , xlab="" , ylab="log2(I)")
p + stat_summary(fun.y=median,geom='point') + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


```{r}
cvSummary <- aggregate(cv ~ condition , data=CVs , median, na.rm=TRUE)
knitr::kable(cvSummary,caption = 'median of cv')
```

\pagebreak

```{r SDViolin, fig.cap="Distribution of protein standard deviation (after sample normalization and scaling) within conditions and overall"}
cond1 <- grp2$getNormalizedConditionData( grp2$conditions[1] )
cond2 <- grp2$getNormalizedConditionData( grp2$conditions[2] )

cond1 <- apply(cond1, 1, sd, na.rm=TRUE)
cond2 <- apply(cond2, 1, sd, na.rm=TRUE)

all <- apply( grp2$getNormalized()$data, 1 , sd, na.rm=TRUE )
SDs<-rbind(data.frame( condition=grp2$conditions[1], sd=cond1), data.frame(condition=grp2$conditions[2],sd=cond2 ), data.frame(condition="all", sd=all))

p <- qplot( condition , sd , data=SDs , geom="violin" , xlab="" , ylab="log2(I)")
p + stat_summary(fun.y=median,geom='point') + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}
sdSummary <-aggregate(sd ~ condition , data=SDs , median, na.rm=TRUE)
knitr::kable(sdSummary, caption = 'median of sd')

progress(0.1, "CVs")

```

\pagebreak

# Heatmaps and Clustering for Samples and Proteins


```{r correlation,dpi=300,fig.height=8,fig.width=8, fig.cap="Heatmap of correlations (spearman) between samples."}
col <- c("red","blue")
simpleheatmap(cor(grp2$getNormalized()$data , use="pairwise.complete.obs", method="spearman")^2,palette = getGreensScale(21),ColSideColors = col[as.factor(grp2$getAnnotation()$Condition)], margins = c(10,3))

```



In Figure \@ref(heatmapData) we show how samples are clustering depending on their protein expression profiles.
Also proteins are clustered according their profiles within the different samples.


```{r heatmapData, fig.width=8, fig.height=10, dpi=300, fig.cap="Heatmap for normalized and scaled values based (log2 transformed data)" }
tmp <- grp2$getNormalized()$data

simpleheatmap(grp2$getNormalized()$data[grp2$getNrNAs() < ncol(grp2$getNormalized()$data)/2,],ColSideColors = col[as.factor(grp2$getAnnotation()$Condition)], margins=c(10,3) )

progress(0.2, "Heatmaps")
```


# FGCZ - Two Group Analysis

Here we show for the normalized matrices the result of a
two group analysis using bioconductor limma. 

```{r include=FALSE}
fixcols <- function(resRel, fcname = "log2(Cond/Cont)", name = "qvalue", top = 20){
  resRel<-do.call("rbind", resRel)
  rownames(resRel) <- NULL
  resRel <- data.frame(resRel,tmp= 2^resRel$foldchange)
  colnames(resRel) <- c("regulation", "Accession", fcname, name, gsub("log2","",fcname))
  
  resRel <- resRel[order(resRel[,name])[1:min(top,nrow(resRel))],]
  
  return(resRel)
}
```

\newpage

## Adjusted moderated p-values (q-values)

Significant calls are made with q Value smaller 0.01 (false discovery rate, computed with bioconductor package qvalue). 

```{r fig.width=10, fig.height=8, dpi=300, fig.cap="VolcanoPlot for normalized and scaled proteins with stringent thresholds. x axis log2 fold change"}
fcname = paste("log2(",grp2$getConditions()$condition,"/",grp2$getConditions()$reference,")",sep="")
res.eb <-grp2$getPValues()
resStr <-quantable::volcanoplot(res.eb$logFC, res.eb$q.mod, foldchangethresh = grp2$qfoldchange , pthresh = grp2$qvalue ,labels=rownames(res.eb),xlab = fcname)

```

\newpage

```{r results="markup"}
if(!is.null(resStr)){
  resStr<-fixcols(resStr,fcname = fcname, name="moderated qvalue")
  knitr::kable(resStr, caption = "Most significant regulated proteins.")
}
```

\newpage

## Moderated p-values

Significant calls are made with *moderated* p Value smaller 0.05. 

```{r fig.width=10, fig.height=8, dpi=300, fig.cap="VolcanoPlot for normalized and imputed matrix (asinh) with moderated p Values."}
res.eb <-grp2$getPValues()
resStr <- quantable::volcanoplot(res.eb$logFC, res.eb$p.mod, foldchangethresh = grp2$pfoldchange , pthresh = grp2$pvalue ,labels=rownames(res.eb), xlab =fcname )
```

\newpage

```{r  results="markup"}
if(!is.null(resStr)){
  resStr<-fixcols(resStr, fcname=fcname , name = "moderated pvalue" )
  knitr::kable(resStr, caption = "Most significant regulated proteins.")
}
```

\newpage

## p-values

Significant calls are made with a p Value smaller 0.05. 


```{r fig.width=10, fig.height=8, dpi=300, fig.cap="VolcanoPlot for normalized and imputed matrix (asinh) with moderated q Values."}
res.eb <-grp2$getPValues()

resRel <- quantable::volcanoplot(res.eb$logFC, res.eb$p.ord, foldchangethresh = grp2$pfoldchange , pthresh = grp2$pvalue ,labels=rownames(res.eb),xlab = fcname)

```

\newpage

```{r  results="markup"}
resRel <- fixcols(resRel,fcname=fcname, name="p-value")
knitr::kable(resRel, caption = "Most significant regulated proteins.")
progress(0.1, "volcanos")
```

## References

The q-values and p-values were computed using the bioconductor package [limma](https://bioconductor.org/packages/release/bioc/html/limma.html).

## Disclaimer and Acknowledgements

This document was generated using Rmarkdown and processes text files which are generated with a label-free quantitation software MaxQuant or Progenesis. 
The obtained results should be validated orthogonally as well (e.g. with Western blots). 


