---
title: "FGCZ Two-Group Analysis for: `r grp2$projectName` \n Statistics for a Quantitative Protein Matrix"
author: "Functional Genomics Center Zurich"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```



# Input Matrix

Experiment is called: `r grp2$experimentName`

## Numbers

The protein matrix is filtered like this:

- Minimum of peptides protein: `r grp2$nrPeptides`
- Maximum of missing values per protein : `r grp2$maxNA`

The total number of proteins in this experiment is: `r nrow(grp2$proteinIntensity)`


```{r }
library(knitr)

tab <- data.frame(table(grp2$annotation$Condition))
colnames(tab) <- c("Condition","# samples")
knitr::kable(tab)


```


# The Groups

Here the files in each group:

```{r }
tab<-grp2$annotation[,c("Condition","Raw.file")]
rownames(tab) <- NULL
knitr::kable(tab)

```

\pagebreak

# Proteins Used for Quantitation


```{r histmissing, }

grp2$proteinIntensity[grp2$proteinIntensity == 0]<-NA
missing <- apply(grp2$proteinIntensity,1,function(x){sum(is.na(x))})
int <- apply(grp2$proteinIntensity,1,sum, na.rm=TRUE)
grp2$proteinIntensity <- grp2$proteinIntensity[order(missing, -int,decreasing = T),]
```


The input matrix has the following structure.


```{r overview, fig.with=9,fig.height=6,fig.cap="Heatmap for quantifyable proteins sorted by missigness and intensity (log2). Gray - missing values"}
library(quantable)

grp2$proteinIntensity[grp2$proteinIntensity == 0]<-NA
z <- log2(grp2$proteinIntensity)
imageWithLabels(t(z), col.labels = NULL, col=quantable::getBlueScale(21))

```

```{r, fig.cap="# of missing values per protein"}

plot(table(missing))
```

\pagebreak


# Distribution of intensities

Shown in Figure \@ref(distributionRaw) are the un-normalized values while in Figure  \@ref(normalized)
are the z transformed values (subtracted median and divided by variance).

```{r distributionRaw, fig.width=7, fig.height=7, fig.cap="Density plot for quantifyable proteins (log2 transformed)"}
library(ggplot2)
library(reshape2)
longm <- melt(z)

p <- qplot( variable , value , data=longm , geom="violin" , xlab="" , ylab="log10(I)")
p + stat_summary(fun.y=median,geom='point') +theme(axis.text.x = element_text(angle = 90, hjust = 1))

```



```{r normalized,  fig.width=7,fig.height=7,dpi=300, fig.cap="Density plot for normalized values (log2)"}
normalized <- quantable::robustscale(z)

longm <- melt(normalized$data)
p <- qplot( variable , value , data=longm , geom="violin" , xlab="" , ylab="log10(I)")
p + stat_summary(fun.y=median,geom='point') + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

The scaling factors are visualized in Figure \@ref(scaling). It shows with what factor the individually samples
are normalized. Large differences in the scaling factors are critical.

```{r  scaling,dpi=300, fig.height=6, fig.cap="Applied scaling factors for normalization (calculated using median normalization)."}
par(mfrow=c(2,1))
par(mar=c(3,10,3,3))

barplot(normalized$medians,horiz=T,las=2, main="median correction", cex.names = 0.6)
barplot(normalized$mads,horiz=T, las=2 , main="standard deviation correction", cex.names = 0.6)

```


\pagebreak

# Coefficients of variation

```{r CViolin, fig.cap="distribution of protein CV within condition and overall"}
cond1 <- grp2$proteinIntensity[, grp2$annotation$Raw.file[grp2$annotation$Condition == grp2$conditions[1]]]
cond2 <- grp2$proteinIntensity[, grp2$annotation$Raw.file[grp2$annotation$Condition == grp2$conditions[2]]]
cond1 <- quantable::CV(cond1)
cond2 <- quantable::CV(cond2)
all <- quantable::CV(grp2$proteinIntensity)
CVs<-rbind(data.frame(condition=grp2$conditions[1], cv=cond1), data.frame(condition=grp2$conditions[2],cv=cond2 ), data.frame(condition="all", cv=all))

p <- qplot( condition , cv , data=CVs , geom="violin" , xlab="" , ylab="log10(I)")
p + stat_summary(fun.y=median,geom='point') + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


```{r}
cvSummary <-aggregate(cv ~ condition , data=CVs , median, na.rm=TRUE)
knitr::kable(cvSummary,caption = 'median of cv')
```

\pagebreak

```{r SDViolin, fig.cap="Distribution of protein SD (after sample normalization) within condition and overall"}
cond1 <- normalized$data[, grp2$annotation$Raw.file[grp2$annotation$Condition == grp2$conditions[1]]]
cond2 <- normalized$data[, grp2$annotation$Raw.file[grp2$annotation$Condition == grp2$conditions[2]]]
cond1 <- apply(cond1,1, sd, na.rm=TRUE)
cond2 <- apply(cond2,1, sd, na.rm=TRUE)
all <- apply(normalized$data,1, sd, na.rm=TRUE)
SDs<-rbind(data.frame(condition=grp2$conditions[1], sd=cond1), data.frame(condition=grp2$conditions[2],sd=cond2 ), data.frame(condition="all", sd=all))

p <- qplot( condition , sd , data=SDs , geom="violin" , xlab="" , ylab="log10(I)")
p + stat_summary(fun.y=median,geom='point') + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}
sdSummary <-aggregate(sd ~ condition , data=SDs , median, na.rm=TRUE)
knitr::kable(sdSummary, caption = 'median of sd')
```

\pagebreak

# Heatmaps and Clustering for Samples and Proteins


```{r correlation,dpi=300,fig.height=8,fig.width=8, fig.cap="Heatmap of correlations (spearman) between samples."}
col <- c("red","blue")
simpleheatmap(cor(normalized$data, use="pairwise.complete.obs", method="spearman")^2,palette = getGreensScale(21),ColSideColors = col[as.factor(grp2$annotation$Condition)], margins = c(10,3))

```



In Figure \@ref(heatmapData) we show how samples are clustering depending on their protein expression profiles.
Also proteins are clustered according their profiles within the different samples.


```{r heatmapData, fig.width=8, fig.height=10, dpi=300 }
nas <-apply(normalized$data,1, function(x){sum(is.na(x))})
simpleheatmap(normalized$data[nas < 6,],ColSideColors = col[as.factor(grp2$annotation$Condition)], margins=c(10,3) )

```


# FGCZ - Two Group Analysis 
