---
title: "FGCZ Two-Group Analysis for: p2084 \n Statistics for a Quantitative Protein Matrix"
author: "Functional Genomics Center Zuerich"
date: "19 July 2016"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r include=FALSE}
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("Figure ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

# Input Matrix

Experiment is called: OID2264

## Numbers

The protein matrix is read like this:

All are QUANTIFIED, as filtering should be done before uploading the matrix file.
Zero values will eventually be imputed depending on the argument.

The total number of proteins in this experiment is: `r nrow(grp2$proteinIntensity)`


```{r }
library(knitr)

tab <- data.frame(table(grp2$annotation$Condition))
colnames(tab) <- c("Condition","# samples")
knitr::kable(tab)


```


# The Groups

Here the files in each group:

```{r }
tab<-grp2$annotation[,c("Condition","Raw.file")]
rownames(tab) <- NULL
knitr::kable(tab)

```

# Proteins Used for Quantitation


```{r histmissing, fig.cap="nr of missing values per protein"}

grp2$proteinIntensity[grp2$proteinIntensity == 0]<-NA
missing <- apply(grp2$proteinIntensity,1,function(x){sum(is.na(x))})
int <- apply(grp2$proteinIntensity,1,sum, na.rm=TRUE)

plot(table(missing))
grp2$proteinIntensity <- grp2$proteinIntensity[order(missing, -int,decreasing = T),]
```


The input matrix has the following structure.


```{r overview, fig.with=9,fig.height=8,fig.cap="Heatmap for quantifyable proteins sorted by missigness and intensity (log2). Gray - missing values"}
library(quantable)
grp2$proteinIntensity[grp2$proteinIntensity == 0]<-NA
z <- log2(grp2$proteinIntensity)
imageWithLabels(t(z), col.labels = NULL, col=quantable::getBlueScale(21))
```


# Distribution of intensities

Shown in Figure \@ref(distributionRaw) are the un-normalized values while in Figure  \@ref(normalized)
are the z transformed values (subtracted median and divided by variance).

```{r distributionRaw, fig.width=7, fig.height=7, fig.cap="Density plot for quantifyable proteins (log2 transformed)"}
library(ggplot2)
library(reshape2)
longm <- melt(z)

p <- qplot( variable , value , data=longm , geom="violin" , xlab="" , ylab="log10(I)")
p + stat_summary(fun.y=median,geom='point') +theme(axis.text.x = element_text(angle = 90, hjust = 1))

```



```{r normalized,  fig.width=7,fig.height=7,dpi=300, fig.cap="Density plot for normalized values (log2)"}
normalized <- quantable::robustscale(z)

longm <- melt(normalized$data)
p <- qplot( variable , value , data=longm , geom="violin" , xlab="" , ylab="log10(I)")
p + stat_summary(fun.y=median,geom='point') + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

The scaling factors are visualized in Figure \@ref(scaling). It shows with what factor the individually samples
are normalized. Large differences in the scaling factors are critical.

```{r  scaling,dpi=300, fig.height=9, fig.cap="Applied scaling factors for normalization (calculated using median normalization)."}
par(mfrow=c(2,1))
par(mar=c(3,10,3,3))

barplot(normalized$medians,horiz=T,las=2, main="median correction", cex.names = 0.6)
barplot(normalized$mads,horiz=T, las=2 , main="standard deviation correction", cex.names = 0.6)

```


```{r correlation,dpi=300,fig.height=8,fig.width=8, fig.cap="Heatmap of correlations (spearman) between samples."}
col <- c("red","blue")
simpleheatmap(cor(normalized$data, use="pairwise.complete.obs", method="spearman")^2,palette = getGreensScale(21),ColSideColors = col[grp2$annotation$Condition], margins = c(10,3))

```

# Heatmaps and Clustering for Proteins

In Figure \@ref(heatmapData) we show how samples are clustering depending on their protein expression profiles.
Also proteins are clustered according their profiles within the different samples.


```{r heatmapData, fig.width=8, fig.height=10, dpi=300 }
nas <-apply(normalized$data,1, function(x){sum(is.na(x))})
simpleheatmap(normalized$data[nas < 6,],ColSideColors = col[grp2$annotation$Condition], margins=c(10,3) )

```


# FGCZ - Two Group Analysis 
