---
title: "Assay Library Report"
author: "Witold Wolski"
mailto: "wew@fgcz.ethz.ch"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
toc : true
---

# Generate Ion Libraries

## Parameters

```{r installLatestversions, echo=FALSE, message=FALSE, eval=TRUE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=10, fig.height=10, warning = FALSE)

library(prozor)
library(specL)

```

```{r prepareEnvLoadLibraries, echo=FALSE, message=FALSE}
R.Version()$version.string
library(specL)
packageVersion('specL')

```

Library generation is run with the following parameters:

```{r setparameters, echo=FALSE}

SWATH_LIBRARY <- file.path(OUTPUTDIR, "assay_library.tsv")
PEPPROTMAPPING <- file.path(OUTPUTDIR, "pepprot.tsv")

FRAGMENTIONMZRANGE <- c(300,1250)
FRAGMENTIONRANGE <- c(MIN_IONS,200)




```

```{r, echo=FALSE}
if(TRUE){
cat(" SWATH_LIBRARY = ", SWATH_LIBRARY, "\n", 
    " PEPPROTMAPPING = ",PEPPROTMAPPING , "\n",
    " NON_REDUNDANT = ",NON_REDUNDANT , "\n",
    " REDUNDANT = ",REDUNDANT, "\n" )
}

cat(" MZ_ERROR = ",MZ_ERROR, "\n",
    " FRAGMENTIONMZRANGE = ", FRAGMENTIONMZRANGE, "\n",
    " FRAGMENTIONRANGE = ",FRAGMENTIONRANGE, "\n",
    " FASTA_FILE = ", FASTA_FILE, "\n",
    " MAX_IONS = ", MAX_IONS, "\n",
    " MIN_IONS = ", MIN_IONS, "\n",
    " MASCOT_MIN_SCORE = ", MASCOT_MIN_SCORE, "\n"
    )

```

```{r readDatabases, eval=TRUE, echo=FALSE, message=FALSE}

print(NON_REDUNDANT)
system.time( nonRedundantBlib <- read.bibliospec(NON_REDUNDANT) )
system.time( redundantBlib <- read.bibliospec(REDUNDANT) )
#save(redundantBlib, file=file.path(OUTPUTDIR,"redundant.Rdata"))
#save(nonRedundantBlib, file=file.path(OUTPUTDIR,"nonredundant.Rdata"))
  
redundantBlibF <- plyr::laply(redundantBlib, function(x){x$mascotScore > MASCOT_MIN_SCORE})
redundantBlib <- redundantBlib[redundantBlibF]

nonRedundantBlibF <- plyr::laply(nonRedundantBlib, function(x){x$mascotScore > MASCOT_MIN_SCORE})
nonRedundantBlib <- nonRedundantBlib[nonRedundantBlibF]
```


Defined filtering function.

```{r defineFragmentFunction}

fragmentIonFunctionUpTo2 <- function (b, y) {
  Hydrogen <- 1.007825
  Oxygen <- 15.994915
  Nitrogen <- 14.003074
  b1_ <- (b )
  y1_ <- (y )
  b2_ <- (b + Hydrogen) / 2
  y2_ <- (y + Hydrogen) / 2 
  return( cbind(b1_, y1_, b2_, y2_) )
}

```


```{r, echo=FALSE, message=FALSE, eval=TRUE}

rtPep <- bibliospec::CiRTpeptides
colnames(rtPep) <- c("peptide", "rt")
specLibrary <- genSwathIonLib(
  data = nonRedundantBlib,
  data.fit = redundantBlib,
  max.mZ.Da.error = MZ_ERROR,
  topN = MAX_IONS,
  fragmentIonMzRange = FRAGMENTIONMZRANGE,
  fragmentIonRange = FRAGMENTIONRANGE,
  fragmentIonFUN = fragmentIonFunctionUpTo2,
  iRT = IRT_PEPTIDES,
  mascotIonScoreCutOFF = IRT_PEPTIDES
  )

op <- par(mfrow=c(2, 1))
class(specLibrary)

plot(specLibrary)

length(nonRedundantBlib)
```

## Library Generation Summary

```{r, echo=FALSE}
length(specLibrary)
slotNames(specLibrary)

length(specLibrary@rt.input)
length(specLibrary@rt.normalized)
specLibrary@ionlibrary[[1]]

slotNames(specLibrary@ionlibrary[[1]])

if(length(specLibrary@ionlibrary) ==0){
  library(knitr)
  opts_chunk$set(eval=FALSE, message=FALSE, echo=FALSE)
}

summary(specLibrary)

```


Total Number of PSM's with Mascot e score < 0.05, in your search is __`r length(redundantBlib)`__. The number of unique precurosors is __`r length(nonRedundantBlib)`__.

The size of the generated ion library is __`r length(specLibrary@ionlibrary)`__.

That means that __`r length(specLibrary@ionlibrary)/length(nonRedundantBlib) *100`__ % of the unique precursors fullfilled the filtering criteria.


# Assigning identified precursors to proteins


```{r getPeptideProt, echo=FALSE, message=FALSE}
protpep = getProteinPeptideTable(specLibrary)
```

```{r prozor, echo=FALSE, message=FALSE }
library(prozor)
fasta = read.fasta(file = FASTA_FILE, as.string = TRUE, seqtype="AA")

protpepDF <- data.frame(protpep)


protpepDF <- plyr::rename(protpepDF, c('peptideSequence'='peptideSeq',
                                       'peptideModSequence' = 'peptideModSeq',
                                       'z' = 'precursorCharge' ))

protpepAnnot = annotatePeptides(protpepDF,fasta)
save(protpepAnnot, file=file.path(OUTPUTDIR,"protprotpepAnnot.rda"))

```

## Annotate peptides

```{r, echo=FALSE, message=FALSE}

library(Matrix)
write.table(protpepAnnot,file=PEPPROTMAPPING,quote = FALSE, row.names = FALSE,sep="\t")
load(file.path(OUTPUTDIR,"protprotpepAnnot.rda"))

pepProtMatrix = prepareMatrix(protpepAnnot, sep=".")
protPepAssingments = greedy(pepProtMatrix)
xx= cbind(names(protPepAssingments),protPepAssingments)


for(i in 1:length(specLibrary@ionlibrary)){
  specl <- specLibrary@ionlibrary[[i]]
  id <- paste(specl@peptideModSeq, specl@prec_z, sep="." )
  tmp = protPepAssingments[[id]]
  if(!is.null(tmp)){
    specLibrary@ionlibrary[[i]]@proteinInformation = tmp
  }
}


uniq <- rowSums(pepProtMatrix)
names(uniq)<-rownames(pepProtMatrix)

proteotyp<- table(uniq)
plot(proteotyp, ylab="nr peptides", xlab="matching number proteins", main="proteotypic")

tp <- unlist(protPepAssingments)

```

Protein FDR `r length(grep("REV_",tp))/length(tp) * 100`


```{r}
freq.table <- table(table(tp))
plot(freq.table,
	ylab = "number of proteins",  
	xlab="number of precurosor assignments", 
	main="single hit wonders")
```

```{r}
freq.table <- table(table(tp))
plot(as.numeric(names(rev(freq.table))), cumsum(rev(freq.table)), 
	type = "h",
	log = "x")

```



- Least specific peptide precursor : `r names(uniq)[which.max(uniq)]` matching `r max(uniq)`

proteins.

- Protein `r names(tmp)[which.max(tmp)]` with most peptide assignments = `r max(tmp)`


Number of annotated precursors is :`r  dim(pepProtMatrix)[1]`

There are in total: `r  dim(protpepAnnot)[1]`
precursor protein assingments.
There are: `r sum(uniq == 1)`

proteotypic precursors, while 
`r dim(pepProtMatrix)[1] - sum(uniq == 1) `
where assigned to 2 or more protein sequences.

## Compute minimal protein set explaining peptides

The  `r dim(pepProtMatrix)[1]` peptide precursors matched protein sequences assigned to 
`r dim(pepProtMatrix)[2]` unique protein identifies.

The minimal protein set explaining precursors has a size of 
`r length(unique(unlist(protPepAssingments)))` 
proteins.

```{r writeSpecnaut, echo=FALSE, message=FALSE}
if(file.exists(SWATH_LIBRARY)){
  file.remove(SWATH_LIBRARY)
}
slotNames(specLibrary)
write.spectronaut(specLibrary,file=SWATH_LIBRARY)
assay_library = read.table(file=SWATH_LIBRARY,header=TRUE,sep="\t")

```

# Session Information

```{r}
sessionInfo()
```

# Summary

The file `assay_library.tsv` contains a spectronaut compatible assay library. The assays in this file are annotated with razor proteins.


# Remarks

This report was generated using the packages:

- [http://bioconductor.org/packages/specL](http://bioconductor.org/packages/specL/) 

- [https://github.com/protViz/prozor](https://github.com/protViz/prozor)


We have invested a lot of time and effort in creating and maintaining this software.
Please cite our publication:

- Panse C, Trachsel C, Grossmann J and Schlapbach R (2015). 
``specL - An R/Bioconductor package to prepare peptide spectrum matches for use in targeted proteomics.'' 
Bioinformatics, pp. 2228-2231. 
[DOI:10.1093/bioinformatics/btv105](http://dx.doi.org/10.1093/bioinformatics/btv105), 
[PMID: 25712692](https://www.ncbi.nlm.nih.gov/pubmed/25712692).

For questions and improvements please do contact the authors of the application [_generateSpecLibrary_](https://github.com/fgcz/generateSpecLibrary/graphs/contributors).

